## Rasterize landcover 

I am classifying landcover using the Alberta Derived Ecosite Phase database [governmentofalberta2020], which classifies landcover polygons according to geography, soil, moisture regime, and dominant vegetation. These instructions are for interacting with the data stored in a PostgreSQL database, but if you have it stored in another place, just modify your code to interact with it that way. 

```{r setup, include=FALSE, cache=FALSE}
#Set root directory to R project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


library(RPostgreSQL)
library(sf)
library(tidyverse)
library(terra)

```


Connect to the PostgreSQL database 
```{r, eval = FALSE} 

# Enter the username and password without putting it into your code
# Note: Once you enter these, they will be visible in your R environment, so I remove them after connecting to the database
username <- rstudioapi::askForPassword("Database username")
password <- rstudioapi::askForPassword("Database password")

pg = dbDriver("PostgreSQL")
# Local Postgres.app database; no password by default
# Of course, you fill in your own database information here.
con = dbConnect(pg, user = username, password = password,
                host = "localhost", port = 5432, dbname = "st_sim_demo")
rm(list = c("username", "password"))

```


Now, to the reason we are using Postgres for some of our data processing. The amount of memory needed to load the dep dataset into R, and then clip it to the L3 FMU layer, would crash the program on most computers. Sending the job to Postgres as a query, though, works just fine. 

```{r, eval = FALSE}

dep_l3 <- st_read(con, query = "SELECT * 
                  FROM dep 
                  WHERE ST_Intersects(dep.shape, (
                  SELECT geometry 
                  FROM fmu_l3
                  WHERE fmu_name = 'L3'));")


dep_l3$raster_code <- as.numeric(as.factor(dep_l3$ep_code)) 

dep_l3 <- st_write(dep_l3, con, "dep_l3")

```

The next step is to rasterize the dep data so that it can be input into the st-sim model. I will start by creating a template raster from the L3 FMU layer using a 100m cell resolution. The I will use the template to rasterize the dep polygons. 
```{r, eval = FALSE}
# Create the template raster 

fmu_l3 <- st_read("0_data/processed/shapefiles/fmu_l3.shp")

box <- st_bbox(fmu_l3)

l3_rast <- terra::crop(rast(xmin = box$xmin, xmax = box$xmax, ymin = box$ymin, ymax = box$ymax, crs = crs(fmu), resolution = 100, vals = 1), vect(fmu_l3), mask = TRUE)

dep_l3_rast <- terra::crop(terra::rasterize(x = vect(dep_l3), y = l3_rast, field = "ep_code"), vect(fmu_l3), mask = TRUE)

# Check to make sure it passes the eye test
plot(dep_l3_rast)

# Write the raster to file
terra::writeRaster(dep_l3_rast, "0_data/processed/rasters/dep_l3.tif", overwrite = TRUE)

```


Finally, I am using a simplified version of the ecosystem classes for this demo using the 9 classes derived from @hart2019, so we need a look up file to reclassify the raster with. 

```{r, eval = FALSE}
# Read in the dep lookup file
dep_lookup <- read.csv("0_data/st-sim/dep_lookup.csv")

# Create the reclassification table 
reclass <- data.frame(id = as.numeric(as.factor(dep_lookup$ep_code)), v = as.numeric(as.factor(dep_lookup$ep_code_hart)))

# Create the new raster 
dep_l3_hart <- dep_l3_rast

# Re-set the ep codes to numeric values (the 'subst' command requires numeric values)
values(dep_l3_hart) <- as.numeric(as.factor(values(dep_l3_hart))) - 1 

# Set the 0 and 41 values to NA
dep_l3_hart <- subst(dep_l3_hart, c(0, 41), NA) 

dep_l3_hart <- subst(x = dep_l3_hart, from = reclass$id, to = reclass$v) 
plot(dep_l3_hart)  # Give it the eye test 


writeRaster(dep_l3_hart, "0_data/processed/rasters/dep_l3_hart.tif", overwrite = TRUE)

```




