## Run basic ST-Sim model 

These are the instructions for creating a SyncroSim library and running a very basic state-and-transition simulation modle using the data created in the previous steps. 

For users who have never worked with SyncroSim before, it would be very helpful to go over the ![getting started gudes and documentation](https://docs.syncrosim.com/)

The bulk of this code was derived from the ![ApexRMS rsyncrosim documentation](https://syncrosim.github.io/rsyncrosim/index.html).

IMPORTANT SETUP INSTRUCTIONS

Before running this script:
  1. Install SyncroSim software (preferably Windows version - see www.syncrosim.com/download)
  2. Install rysncrosim, raster and rgdal R packages (from CRAN)

Note that this Exercise was developed against the following:

* SyncroSim - version 2.4.18 (note that instructions below assume Windows version but works also with Linux)  
* R - version 4.2.2  
* SyncroSim packages:  
  * stim - version 3.3.10  
* R packages:  
  *  rsyncrosim - version 1.4.2
  * terra - version 1.7-37  (which requires rgdal package to be installed also)
  * this.path - version 1.2.0


### Task 1: Setup
```{r setup, cache=FALSE, eval = FALSE}
#Set root directory to R project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

# Load R packages
library(rsyncrosim) # package for working with SyncroSim
library(terra)      # package for working with raster data
library(this.path)  # package for setting the working directory
library(tidyverse)  # For easier data manipulation
library(tidyterra)  


# Check to see if the stsim SyncroSim package is installed (and install it if necessary)
myInstalledPackages = package()
if (!(is.element("stsim", myInstalledPackages$name))) addPackage("stsim")


```

### Task 2: Create the new library
```{r create library, eval = FALSE}
# Create a new library
# NOTE: this will only create a new library if the file doesn't exist already
myLibrary <- ssimLibrary("2_pipeline/st-sim/alpac_l3_demo")


# Set up the Project Definitions

myProject <- rsyncrosim::project(myLibrary, project="Definitions")
rsyncrosim::project(myLibrary, summary=TRUE)     # Lists the projects in this library

# Display internal names of all the project's datasheets - corresponds to the Project Properties in SyncroSim
dataSheets <- datasheet(myProject, summary=T)
```

Re-set some of the terminology 
```{r terminology, eval = FALSE}
# Terminology: change units to 'Hecatares' and labels to 'Forest Type'
sheetData <- datasheet(myProject, "stsim_Terminology")
sheetData
sheetData$AmountUnits[1] <- "Hectares"
sheetData$StateLabelX[1] <- "Forest Type"
saveDatasheet(myProject, sheetData, "stsim_Terminology")
datasheet(myProject, "stsim_Terminology")
```

Define the strata and state classes 
```{r strata-classes, eval = FALSE}
# Stratum: make the primary stratum 'Entire Forest'
sheetData <- datasheet(myProject, "stsim_Stratum", optional = TRUE, empty=T)   # Returns empty dataframe with only required column(s)
sheetData <- addRow(sheetData, data.frame(Name = "Entire Forest", ID = 1))
saveDatasheet(myProject, sheetData, "stsim_Stratum", force=T)
datasheet(myProject, "stsim_Stratum", optional=T)   # Returns entire dataframe, including optional columns

# First State Class Label (i.e. Forest Types): get from the 
dep_lookup <- read.csv("0_data/st-sim/dep_lookup.csv")

states <- data.frame(dep_lookup %>% distinct(ep_code_hart, hab_type_hart))

saveDatasheet(myProject, data.frame(Name = states$ep_code_hart, Description = states$hab_type_hart), "stsim_StateLabelX", force=T)
datasheet(myProject, "stsim_StateLabelX", optional = TRUE)

# Second State Label 
saveDatasheet(myProject, data.frame(Name="All"), "stsim_StateLabelY", force=T)
datasheet(myProject, "stsim_StateLabelY", optional = TRUE)
```

List the transition types to be used in the model 
```{r transition-types, eval = FALSE}
# Transition Types
transitionTypes <- data.frame(Name = "Succession", ID = 1, Color = "255,0,128,0")
saveDatasheet(myProject, transitionTypes, "stsim_TransitionType", force=T)
datasheet(myProject, "stsim_TransitionType", optional=T)   # Returns entire dataframe, including optional columns
```

Create the final state classes for the model 
```{r state classes, eval = FALSE}
# State Classes
stateClasses <- datasheet(myProject, name="stsim_StateClass", empty = TRUE, optional = TRUE)

for(i in 1:nrow(states)){
  stateClasses <- addRow(stateClasses, data.frame(Name = paste0(states$ep_code_hart[i], ":All"), 
                                                  StateLabelXID = states$ep_code_hart[i], 
                                                  StateLabelYID = "All", 
                                                  ID = i))
}
saveDatasheet(myProject, stateClasses, "stsim_StateClass", force=T)
datasheet(myProject, "stsim_StateClass", optional=T)
```

### Task 3: Create new Scenario (within this Project) and setup the Scenario Properties

```{r scenario, eval = FALSE}

# Create a new SyncroSim "Succession Only" scenario
myScenario <- scenario(myProject, "Succession Only")

# Display the internal names of all the scenario datasheets
myDataSheetGuide <- subset(datasheet(myScenario, summary=T), scope == "scenario")   # Generate list of all Datasheets as reference

# Edit the scenario datasheets:

# Run Control - Note that we will set this as a non-spatial run
sheetName <- "stsim_RunControl"
sheetData <- data.frame(MaximumIteration = 5, MinimumTimestep = 0, MaximumTimestep = 10, isSpatial = T)
saveDatasheet(myScenario, sheetData, sheetName)
datasheet(myScenario, "stsim_RunControl")

```

Set the transition probabilities 
```{r transition probabilities, eval = FALSE}

# Deterministic transitions 
sheetName <- "stsim_DeterministicTransition"
sheetData <- datasheet(myScenario, sheetName, empty=T)

trans <- stateClasses %>% filter(Name != "f:All") %>% select(Name) 
trans$Location <- sapply(1:nrow(trans), function(x) paste0("A", x))
colnames(trans) <- colnames(sheetData)

for(i in 1:nrow(trans)){
  sheetData <- addRow(sheetData, data.frame(trans[i, ]))
}
saveDatasheet(myScenario, sheetData, sheetName)
datasheet(myScenario, "stsim_DeterministicTransition")

# Get the transition probabilities datasheet 
transitions <- read.csv("0_data/st-sim/hart-transition-probs.csv")

# Add the probabilities during each time step 
sheetName <- "stsim_Transition"
sheetData <- datasheet(myScenario, sheetName, optional=T, empty=T) 
sheetData <- sheetData %>% select(StateClassIDSource, StateClassIDDest, TransitionTypeID, Probability, AgeMin)

# Format the data
trans_2 <- data.frame(StateClassIDSource = stateClasses$Name[match(transitions$From_class, stateClasses$StateLabelXID)], 
                      StateClassIDDest = stateClasses$Name[match(transitions$To_class, stateClasses$StateLabelXID)], 
                      TransitionTypeID = transitions$Type, 
                      Probability = transitions$Probability, 
                      AgeMin = transitions$Min_age)
trans_2[trans_2 == 0] <- NA

# use the addRow function to add each transition 
for(i in 1:nrow(trans_2)){
  sheetData <- addRow(sheetData, trans_2[i, ])
}
saveDatasheet(myScenario, sheetData, sheetName)
datasheet(myScenario, "stsim_Transition")

```

Set the initial conditions 
```{r initial conditions, eval = FALSE}

# Initial Conditions (spatial) 
sheetName <- "stsim_InitialConditionsSpatial"
sheetData <- datasheet(myScenario, sheetName , optional=T, empty=T)
fp <- file.path(getwd(), "0_data/processed/rasters")

sheetData <- addRow(sheetData, data.frame(StratumFileName = file.path(fp, "l3_rast.tif"), StateClassFileName = file.path(fp, "dep_l3_hart.tif"), 
                                          AgeFileName = file.path(fp, "age_l3_sim.tif")))

saveDatasheet(myScenario, sheetData, sheetName)
datasheet(myScenario, "stsim_InitialConditionsSpatial")


```

```{r output options, eval = FALSE}

# Set the tabular output options 
saveDatasheet(myScenario, data.frame(SummaryOutputSC = TRUE, SummaryOutputSCTimesteps = 1, SummaryOutputSCAges = TRUE, stsim_OutputSpatialAverageAge = TRUE), "stsim_OutputOptions")
              
datasheet(myScenario, "stsim_OutputOptions")


# Set the spatial output options 
saveDatasheet(myScenario, data.frame(RasterOutputSC = TRUE, RasterOutputSCTimesteps = 1, 
                                     RasterOutputAge = TRUE, RasterOutputAgeTimesteps = 1, 
                                     RasterOutputTR = TRUE, RasterOutputTRTimesteps = 1), 
              "stsim_OutputOptionsSpatial")
datasheet(myScenario, "stsim_OutputOptionsSpatial")

```


### Task 4: Run the scenario and analyze results
```{r run, eval = FALSE} 
# Run the scenario: 
resultSummary <- run(myProject, scenario="Succession Only", jobs=6)   # Uses multiprocessing
resultSummary

myDataSheetGuide <- datasheet(myScenario, summary = TRUE)  # The list is long! 

backup(myLibrary)  # Backup of your library - automatically zipped into a .backup subfolder


```

Analyze the tabular output 
```{r analyze, eval = FALSE}

scenario(myLibrary)
myResults <- scenario(myLibrary, scenario = 2)

# Retrieve raw tabular state class tabular output into a dataframe (see myDatasheetGuide for valid names)
outRaw <- datasheet(myResults, name="stsim_OutputStratumState") %>% dplyr::arrange(Iteration, StateLabelXID, Timestep)

# Show a bit of this dataframe
# It is raw output, so it has lots of rows - however you can view it also using RStudio "Environment" pane
tail(outRaw)

## Re-do the previous aggregation and plotting using dplyr and ggplot2
outSum <- data.frame(outRaw %>% group_by(Iteration, Timestep, StateLabelXID) %>% summarise(Amount = sum(Amount))) 

outSummary <- data.frame(outSum %>% group_by(Timestep, StateLabelXID) %>% summarise(Mean = mean(Amount), Minimum = min(Amount), Maximum = max(Amount))) 

ggplot() + 
  geom_line(data = outSummary %>% filter(StateLabelXID == "b2"), colour = "blue",
            aes(x = Timestep, y = Mean)) + 
  geom_ribbon(data = outSummary %>% filter(StateLabelXID == "b2"), aes(0:50, ymin = Minimum, ymax = Maximum), fill = "blue", alpha = 0.1) +
  ylab("Total mixed forest area (ha)")


```

```{r, eval = FALSE}
# Retrieve raw tabular state class tabular output into a dataframe (see myDatasheetGuide for valid names)
outRaw_age <- datasheet(myResults, name="stsim_OutputSpatialAverageAge") 

%>% dplyr::arrange(Iteration, StateLabelXID, Timestep)

```






