--- 
title: "Demonstration of running a biodiversity risk assessment with SyncroSim software, and the ST-Sim package, via R"
author: "Andy Crosby"
site: bookdown::bookdown_site
output: 
  md_document:
    variant: gfm
    toc: TRUE
    toc_depth: 2
    number_sections: TRUE
#bibliography: library.bib
#citation-style: ../../MendeleyBib/CitationStyles/nature.csl
#citation-style: https://www.zotero.org/styles/elsevier-harvard
documentclass: book
link-citations: true
github-repo: https://github.com/alberta-conservation/st-sim-risk-assessment-demo.git
description: "This is document walks through the process of running a basic state-and-transition simulation model and assessing population risk to a species of interest."


---

# Overview
One of the most pressing problems in ecology and conservation is understanding effects of anthropogenic disturbance and climate change on wildlife populations and biodiversity [@Venier2021]. Complicating this understanding is the fact that these cumulative effects are occurring within the context of global climate change, meaning that models based on historical responses to disturbance might no longer be valid [@Jackson2021], leading to unanticipated outcomes from management actions. Anticipating and responding to potential ecosystem changes from the interactions of anthropogenic disturbance with climate change, and risks to biodiversity from those changes, requires a probabilistic ecological forecasting approach within a risk assessment framework to quantify the risks and uncertainties, and ultimately inform the decision-making process [@milner-gulland2010a].  



**Project description**
The Objective of this project is to develop a simple state-and-transition simulation model and use it to assess the forecast population responses by a species of interest. 




```{r GlobalOptions, echo=FALSE}
options(bookdown.render.file_scope = FALSE, knitr.duplicate.label = "allow") 
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r setup, include=FALSE, cache=FALSE}
#Set root directory to R project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r echo=FALSE, out.width = '100%'}
library(kableExtra)
library(knitr)
```


# Importing a geodatabase into PostgreSQL 
```{r child = '1_scripts/00_importing-geodatabase-to-postgres.Rmd'}
```

# Data processing  
```{r child = '1_scripts/01-1_rasterize-subregions.Rmd'}
```

```{r child = '1_scripts/01-2_rasterize-dep.Rmd'}
```

```{r child = '1_scripts/01-3_historic-fire-distributions.Rmd'}
```

```{r child = '1_scripts/01-4_rasterize-seismic.Rmd'}
```

# Run ST-Sim models through SyncroSim software 
```{r child = '1_scripts/02-1_run-basic-stsim.Rmd'}
```



# References


<div id="refs"></div>
